---
title: "COVID-19: First and Second Digits Analysis"
author: "David Jackson"
date: "01/31/2021 3:00 PM/EST"
output: 
  html_document: 
    fig_width: 9
    fig_height: 7
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE,message = FALSE)
library(ggplot2)
library(scales)
library(RSQLite)
library(dplyr)
library(forecast)
library(BenfordTests)
library(benford.analysis)
library(plotly)
```

## Benford's Law Analysis for New Yoirk and Selected US States (Cases and Deaths)
### Is New York State "Cooking their Books" for Covid19 Deaths?
#### Comparing New  York State to:
* Texas,Florida,Illinois,Pennsylvania,Ohio,California,Georgia
  

#### New York Times (Jan. 28, 2020)

* N.Y. Severely Undercounted Virus Deaths in Nursing Homes
"The state attorney general, Letitia James, said it’s likely that the Cuomo administration failed to report thousands of Covid-19 deaths of nursing home residents."
* Source: https://www.nytimes.com/2021/01/28/nyregion/nursing-home-deaths-cuomo.html

#### New York Times COVID-19 Github Repository

* NEW-YORK-DEATHS-METHODOLOGY.md
"Starting with data for Aug. 6, we are switching to reporting deaths in New York State by county of residence instead of the county in which the death took place."
* Sourcs: https://github.com/nytimes/covid-19-data/blob/master/NEW-YORK-DEATHS-METHODOLOGY.md
* Data Source: https://github.com/nytimes/covid-19-data

### Introduction:
* If based on the two quotes from the New York Times that New York State is intentionally understating the deaths in Nursing Homes.
* Analysis will be done using Benford's Law first (1st) digit distribution.
* The conclusion,  Any state  that's first digit analysis (Covid-19 Deaths) is simalar to New York State deaths is probably also "fiddling" with their deaths counts.


### Data Visualization for Covid-19 Daily Cases and Deaths


```{r}
rm(list=ls())
db <- db <- dbConnect(RSQLite::SQLite(),dbname= "../COVID-19-DB/NYTimes.sqlite3")
df <- dbGetQuery(db,"select date,state,county,new_cases,new_deaths from COUNTYDAILY")
df <- subset(df,state %in% c("New York","Texas","Florida","Illinois",
                            "Pennsylvania","Ohio","California",
                             "Georgia" ))
df$date <- as.Date(df$date,format="%Y-%m-%d")
```
#### Plot of Daily Cases by State

```{r}
ggplot(df) + geom_line(aes(x=date,y=new_cases,col=state)) +
  labs(title = "Daily Cases by State") +ylim(0,17000) 

df %>% filter(new_cases >=0) %>%
 ggplot() + geom_line(aes(x=date,y=new_cases,col=state)) +
  labs(title = "Daily Deaths by State")  + facet_wrap(~state,ncol=2,scale="free_y") 
   
```

#### Plot of Daily Deaths by State

```{r}
ggplot(df) + geom_area(aes(x=date,y=new_deaths,col=state)) +
  labs(title = "Daily Deaths by State") +scale_y_log10() + 
  scale_y_continuous(labels=comma) + ylim(0, 2500)

df %>% filter(new_deaths >=0) %>%
 ggplot() + geom_line(aes(x=date,y=new_deaths,col=state)) +
  labs(title = "Daily Deaths by State")  + facet_wrap(~state,ncol=2,scale="free_y") 
   
```

### Benford Analysis of Daily  Cases by State
### New York State

```{r}

set.seed(1234)
nys_cases <- df %>% filter(state =="New York") %>% select(new_cases)
nys_cases <- nys_cases$new_cases
signifd.analysis(nys_cases,digits = 1,ci_col = "lightgreen")
bfd <-benford(nys_cases)
```

####  Mean Absolute Deviation (MAD)

```{r}
MAD(bfd)
```

### California


```{r}
set.seed(1234)
tx_cases <- df %>% filter(state =="California") %>% select(new_cases)
tx_cases <- tx_cases$new_cases
signifd.analysis(tx_cases,digits = 1,ci_col = "lightgreen")
bfd <-benford(tx_cases)
```

*  Mean Absolute Deviation (MAD)

```{r}
MAD(bfd)
```


### Texas

```{r}
set.seed(1234)
tx_cases <- df %>% filter(state =="Texas") %>% select(new_cases)
tx_cases <- tx_cases$new_cases
signifd.analysis(tx_cases,digits = 1,ci_col = "lightgreen")
bfd <-benford(tx_cases)

```
*  Mean Absolute Deviation (MAD)

```{r}
MAD(bfd)
```

### Florida

```{r}
set.seed(1234)
fl_cases <- df %>% filter(state =="Florida") %>% select(new_cases)
fl_cases <- fl_cases$new_cases
signifd.analysis(fl_cases,digits = 1,ci_col = "lightgreen")
bfd <-benford(fl_cases)
MAD(bfd)
```

*  Mean Absolute Deviation (MAD)

```{r}
MAD(bfd)
```

### Illinois

```{r}
set.seed(1234)
il_cases <- df %>% filter(state =="Illinois") %>% select(new_cases)
il_cases <- il_cases$new_cases
signifd.analysis(il_cases,digits = 1,ci_col = "lightgreen")
bfd <-benford(fl_cases)
```

*  Mean Absolute Deviation (MAD)

```{r}
MAD(bfd)
```

### Pennsylvania

```{r}
set.seed(1234)
penn_cases <- df %>% filter(state =="Pennsylvania") %>% select(new_cases)
penn_cases <- penn_cases$new_cases
signifd.analysis(penn_cases,digits = 1,ci_col = "lightgreen")
bfd <-benford(penn_cases)
```

*  Mean Absolute Deviation (MAD)

```{r}
MAD(bfd)
```

### Ohio

```{r}
set.seed(1234)
penn_cases <- df %>% filter(state =="Ohio") %>% select(new_cases)
penn_cases <- penn_cases$new_cases
signifd.analysis(penn_cases,digits = 1,ci_col = "lightgreen")
bfd <-benford(penn_cases)
```

*  Mean Absolute Deviation (MAD)

```{r}
MAD(bfd)
```

### California

```{r}
set.seed(1234)
ca_cases <- df %>% filter(state =="California") %>% select(new_cases)
ca_cases <- ca_cases$new_cases
signifd.analysis(ca_cases,digits = 1,ci_col = "lightgreen")
bfd <-benford(ca_cases)
```

*  Mean Absolute Deviation (MAD)

```{r}
MAD(bfd)
```
