---
title: "Comparting New York With Other States"
author: "David Jackson"
date: "01/30/2021 9:00 PM/EST"
output: 
  html_document: 
    fig_width: 9
    fig_height: 7
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE,message = FALSE)
library(ggplot2)
library(scales)
library(RSQLite)
library(dplyr)
library(forecast)
library(BenfordTests)
library(benford.analysis)
```

### Import New State County Data

```{r}
rm(list=ls())
db <- db <- dbConnect(RSQLite::SQLite(),dbname= "../COVID-19-DB/NYTimes.sqlite3")
df <- dbGetQuery(db,"select date,state,county,new_cases,new_deaths from COUNTYDAILY")
df <- subset(df,state %in% c("New York","Texas","Florida","Illinois","Pennsylvania","Ohio","California",
  "Georgia","North Carolina","New Jersey","Virginia","Washington","Arizona" ))
df$date <- as.Date(df$date,format="%Y-%m-%d")
summary(df$new_cases,na.rm=TRUE)
summary(df$new_deaths,na.rm=TRUE)
```

```{r}
state_totals <- df %>% group_by(state) %>% summarise(
                Cases = sum(new_cases),
                Deaths = sum(new_deaths)
)
daily_totals <- df %>% group_by(date) %>% summarise(
                Cases = sum(new_cases),
                Deaths = sum(new_deaths)
)
```

### Plot of Daily Cases by State

```{r}
ggplot(df) + geom_line(aes(x=date,y=new_cases,col=state)) +
  labs(title = "Daily Cases by State") +ylim(0,17000) 

df %>% filter(new_cases >=0) %>%
 ggplot() + geom_line(aes(x=date,y=new_cases,col=state)) +
  labs(title = "Daily Deaths by State")  + facet_wrap(~state,ncol=2,scale="free_y") 
   
```

### Plot of Daily Cases and Deaths by State

```{r}
ggplot(df) + geom_area(aes(x=date,y=new_cases,col=state)) +
  labs(title = "Daily Cases by State") +scale_y_log10() + 
  scale_y_continuous(labels=comma)

df %>% filter(new_deaths >=0) %>%
 ggplot() + geom_line(aes(x=date,y=new_cases,col=state)) +
  labs(title = "Daily Cases by State")  + facet_wrap(~state,ncol=2,scale="free_y") 

  
ggplot(df) + geom_area(aes(x=date,y=new_deaths,col=state)) +
  labs(title = "Daily Deaths by State") +scale_y_log10() + 
  scale_y_continuous(labels=comma)

df %>% filter(new_deaths >=0) %>%
 ggplot() + geom_line(aes(x=date,y=new_deaths,col=state)) +
  labs(title = "Daily Deaths by State")  + facet_wrap(~state,ncol=2,scale="free_y") 
    
```

## Benford Law for Cases: Selected States


### Benford Analysis of Daily  Cases by State
#### New York State

```{r}
dat <- dbGetQuery(db,"select date,state,county,new_cases,new_deaths from COUNTYDAILY")
```
```{r}

set.seed(1234)
nys_cases <- df %>% filter(state =="New York") %>% select(new_cases)
nys_cases <- nys_cases$new_cases
signifd.analysis(nys_cases,digits = 1,ci_col = "lightgreen")
bfd <-benford(nys_cases)
MAD(bfd)
```

#### Texas

```{r}
set.seed(1234)
tx_cases <- df %>% filter(state =="Texas") %>% select(new_cases)
tx_cases <- tx_cases$new_cases
signifd.analysis(tx_cases,digits = 1,ci_col = "lightgreen")
bfd <-benford(tx_cases)
MAD(bfd)
```

#### Florida

```{r}
set.seed(1234)
fl_cases <- df %>% filter(state =="Florida") %>% select(new_cases)
fl_cases <- fl_cases$new_cases
signifd.analysis(fl_cases,digits = 1,ci_col = "lightgreen")
bfd <-benford(fl_cases)
MAD(bfd)
```

#### Illinois

```{r}
set.seed(1234)
il_cases <- df %>% filter(state =="Illinois") %>% select(new_cases)
il_cases <- il_cases$new_cases
signifd.analysis(il_cases,digits = 1,ci_col = "lightgreen")
bfd <-benford(fl_cases)
MAD(bfd)
```

#### Pennsylvania

```{r}
set.seed(1234)
penn_cases <- df %>% filter(state =="Pennsylvania") %>% select(new_cases)
penn_cases <- penn_cases$new_cases
signifd.analysis(penn_cases,digits = 1,ci_col = "lightgreen")
bfd <-benford(penn_cases)
MAD(bfd)
```

#### Ohio

```{r}
set.seed(1234)
penn_cases <- df %>% filter(state =="Ohio") %>% select(new_cases)
penn_cases <- penn_cases$new_cases
signifd.analysis(penn_cases,digits = 1,ci_col = "lightgreen")
bfd <-benford(penn_cases)
MAD(bfd)
```

#### California

```{r}
set.seed(1234)
ca_cases <- df %>% filter(state =="California") %>% select(new_cases)
ca_cases <- ca_cases$new_cases
signifd.analysis(ca_cases,digits = 1,ci_col = "lightgreen")
bfd <-benford(ca_cases)
MAD(bfd)
```

#### Georgia

```{r}
set.seed(1234)
ga_cases <- df %>% filter(state =="Georgia") %>% select(new_cases)
ga_cases <- ga_cases$new_cases
signifd.analysis(ga_cases,digits = 1,ci_col = "lightgreen")
bfd <-benford(ga_cases)
MAD(bfd)
```


#### North Carolina

```{r}
set.seed(1234)
nc_cases <- dat %>% filter(state =="North Carolina") %>% select(new_cases)
nc_cases <- nc_cases$new_cases
signifd.analysis(nc_cases,digits = 1,ci_col = "lightgreen")
bfd <-benford(nc_cases)
MAD(bfd)
```

#### Michigan

```{r}
set.seed(1234)
nc_cases <- dat %>% filter(state =="Michigan") %>% select(new_cases)
nc_cases <- nc_cases$new_cases
signifd.analysis(nc_cases,digits = 1,ci_col = "lightgreen")
bfd <-benford(nc_cases)
MAD(bfd)
```

#### New Jersey

```{r}
set.seed(1234)
nc_cases <- dat %>% filter(state =="New Jersey") %>% select(new_cases)
nc_cases <- nc_cases$new_cases
signifd.analysis(nc_cases,digits = 1,ci_col = "lightgreen")
bfd <-benford(nc_cases)
MAD(bfd)
```

#### Virginia

```{r}
set.seed(1234)
vi_cases <- dat %>% filter(state =="Virginia") %>% select(new_cases)
vi_cases <- vi_cases$new_cases
signifd.analysis(vi_cases,digits = 1,ci_col = "lightgreen")
bfd <-benford(vi_cases)
MAD(bfd)
```

#### Washington

```{r}
set.seed(1234)
wa_cases <- dat %>% filter(state =="Washington") %>% select(new_cases)
wa_cases <- wa_cases$new_cases
signifd.analysis(wa_cases,digits = 1,ci_col = "lightgreen")
bfd <-benford(vi_cases)
MAD(bfd)
```

#### Arizona

```{r}
set.seed(1234)
wa_cases <- dat %>% filter(state =="Arizona") %>% select(new_cases)
wa_cases <- wa_cases$new_cases
signifd.analysis(wa_cases,digits = 1,ci_col = "lightgreen")
bfd <-benford(vi_cases)
MAD(bfd)
```


### Benford Law for Deaths: Selected States

#### New York State

```{r}
set.seed(1234)
nys_cases <- df %>% filter(state =="New York") %>% select(new_deaths)
nys_cases <- nys_cases$new_deaths
signifd.analysis(nys_cases,digits = 1,ci_col = "lightgreen")
bfd <-benford(nys_cases)
MAD(bfd)
```

#### Texas

```{r}
set.seed(1234)
tx_cases <- df %>% filter(state =="Texas") %>% select(new_deaths)
tx_cases <- tx_cases$new_deaths
signifd.analysis(tx_cases,digits = 1,ci_col = "lightgreen")
bfd <-benford(tx_cases)
MAD(bfd)
```

#### Florida

```{r}
set.seed(1234)
fl_cases <- df %>% filter(state =="Florida") %>% select(new_deaths)
fl_cases <- fl_cases$new_deaths
signifd.analysis(fl_cases,digits = 1,ci_col = "lightgreen")
bfd <-benford(fl_cases)
MAD(bfd)
```

#### Illinois

```{r}
set.seed(1234)
il_cases <- df %>% filter(state =="Illinois") %>% select(new_deaths)
il_cases <- il_cases$new_deaths
signifd.analysis(il_cases,digits = 1,ci_col = "lightgreen")
bfd <-benford(fl_cases)
MAD(bfd)
```

#### Pennsylvania

```{r}
set.seed(1234)
penn_cases <- df %>% filter(state =="Pennsylvania") %>% select(new_deaths)
penn_cases <- penn_cases$new_deaths
signifd.analysis(penn_cases,digits = 1,ci_col = "lightgreen")
bfd <-benford(penn_cases)
MAD(bfd)
```

#### Ohio

```{r}
set.seed(1234)
penn_cases <- df %>% filter(state =="Ohio") %>% select(new_deaths)
penn_cases <- penn_cases$new_deaths
signifd.analysis(penn_cases,digits = 1,ci_col = "lightgreen")
bfd <-benford(penn_cases)
MAD(bfd)
```

#### California

```{r}
set.seed(1234)
penn_cases <- df %>% filter(state =="California") %>% select(new_deaths)
penn_cases <- penn_cases$new_deaths
signifd.analysis(penn_cases,digits = 1,ci_col = "lightgreen")
bfd <-benford(penn_cases)
MAD(bfd)
```

#### Georgia

```{r}
set.seed(1234)
ga_cases <- df %>% filter(state =="Georgia") %>% select(new_deaths)
ga_cases <- ga_cases$new_deaths
signifd.analysis(ga_cases,digits = 1,ci_col = "lightgreen")
bfd <-benford(ga_cases)
MAD(bfd)
```


#### North Carolina

```{r}
set.seed(1234)
ga_cases <- dat %>% filter(state =="North Carolina") %>% select(new_deaths)
ga_cases <- ga_cases$new_deaths
signifd.analysis(ga_cases,digits = 1,ci_col = "lightgreen")
bfd <-benford(ga_cases)
MAD(bfd)
```

#### Michigan

```{r}
set.seed(1234)
ga_cases <- dat %>% filter(state =="Michigan") %>% select(new_deaths)
ga_cases <- ga_cases$new_deaths
signifd.analysis(ga_cases,digits = 1,ci_col = "lightgreen")
bfd <-benford(ga_cases)
MAD(bfd)
```

#### New Jersey

```{r}
set.seed(1234)
ga_cases <- dat %>% filter(state =="New Jersey") %>% select(new_deaths)
ga_cases <- ga_cases$new_deaths
signifd.analysis(ga_cases,digits = 1,ci_col = "lightgreen")
bfd <-benford(ga_cases)
MAD(bfd)
```

#### Virginia

```{r}
set.seed(1234)
vi_cases <- dat %>% filter(state =="Virginia") %>% select(new_deaths)
vi_cases <- vi_cases$new_deaths
signifd.analysis(vi_cases,digits = 1,ci_col = "lightgreen")
bfd <-benford(vi_cases)
MAD(bfd)
```


#### Washington

```{r}
set.seed(1234)
wa_cases <- dat %>% filter(state =="Washington") %>% select(new_deaths)
wa_cases <- wa_cases$new_deaths
signifd.analysis(wa_cases,digits = 1,ci_col = "lightgreen")
bfd <-benford(wa_cases)
MAD(bfd)
```

#### Arizona

```{r}
set.seed(1234)
wa_cases <- dat %>% filter(state =="Arizona") %>% select(new_deaths)
wa_cases <- wa_cases$new_deaths
signifd.analysis(wa_cases,digits = 1,ci_col = "lightgreen")
bfd <-benford(wa_cases)
MAD(bfd)
```