---
title: "New York's COVID-19"
author: "David Jackson"
date: "Nov. 26, 2020"
output: 
  html_document: 
    fig_width: 9
    fig_height: 7
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,message = FALSE,warning = FALSE)
library(ggplot2)
library(scales)
library(RSQLite)
library(dplyr)
theme_set(theme_linedraw() + theme(panel.grid = element_line(linetype = 'dashed')))
```

## New York's Battle with COVID-19
Starting with data for Aug. 6, we are switching to reporting deaths in New York State by county of residence instead of the county in which the death took place. (Source: New York Times)
```{r}
rm(list=ls())
source("../DATA/movavg.R")
db <- db <- dbConnect(RSQLite::SQLite(),dbname= "../COVID-19-DB/NYTimes.sqlite3")
df <- dbGetQuery(db,"select date,state,new_cases,new_deaths from STATESDAILY")
df <- subset(df,state=="New York"  )
df$date <- as.Date(df$date)
mean(df$new_cases)
mean(df$new_deaths)
```
```{r}
db <- db <- dbConnect(RSQLite::SQLite(),dbname= "../COVID-19-DB/NYTimes.sqlite3")
df1<- dbGetQuery(db,"select date,state,county,new_cases,new_deaths from COUNTYDAILY")
CA <- subset(df1,state=="New York" & county=="New York City"  )
CA$date <- as.Date(CA$date)
CA <- subset(CA,date >="2020-09-01")
mean(CA$new_cases)
mean(CA$new_deaths)
NYT <- subset(df1,state=="New York")
```
```{r}
CA %>% group_by(county) %>% summarize(
  Cases = sum(new_cases,na.rm = TRUE),
  Deaths = sum(new_deaths,na.rm =TRUE)
)
```


```{r}
US <- subset(df,date >="2020-03-03")

US <- US[order(US$date),]
US$MA14 <- movingAverage(US$new_cases,14)
US$MAD <- movingAverage(US$new_deaths,14)
```

### Plot of Daily new_cases and new_deaths

```{r}
ggplot(US) + geom_line(aes(x=date,y=new_cases,col="Daily new_cases")) +
  labs(title="COVID-19 new_cases by Date") +
  geom_smooth(aes(x=date,y=new_cases,col="Loess"),span=0.25) +
   scale_x_date()
  
```
```{r}
ggplot(US) + geom_line(aes(x=date,y=new_deaths,col="Daily new_deaths")) +
  labs(title="COVID-19 new_deaths by Date") + ylim(0,175) +
  geom_smooth(aes(x=date,y=new_deaths,col="Loess"),span=0.25) +
   scale_x_date()
```
```{r}
USA <- subset(US,date >="2020-06-01")
ggplot(USA) + geom_line(aes(x=date,y=new_cases,col="Daily new_cases")) +
  labs(title="COVID-19 new_cases by Date (since Jun. 1, 2020)") +
  geom_smooth(aes(x=date,y=new_cases,col="Loess"),span=0.25) +
   scale_x_date()

ggplot(USA) + geom_line(aes(x=date,y=new_deaths,col="Daily new_deaths")) +
  labs(title="COVID-19 new_deaths by Date (since Jun. 1, 2020)") + ylim(0,175) +
  geom_smooth(aes(x=date,y=new_deaths,col="Loess"),span=0.25) +
   scale_x_date()
```

#### Non-Moving Average By Week and By Month

```{r}
US$Monthly <- as.Date(cut(US$date,
  breaks = "month"))
US$Weekly <- as.Date(cut(US$date,
  breaks = "week",
  start.on.monday = FALSE))

```
```{r}
Weekly_new_cases <- aggregate(new_cases~Weekly,US,FUN=sum)
Weekly_new_deaths <- aggregate(new_deaths~Weekly,US,FUN=sum)
Weekly_new_cases$DRate <- Weekly_new_deaths$new_deaths/Weekly_new_cases$new_cases
Weekly_new_cases$LivedSaved <- Weekly_new_cases$new_cases * (max(Weekly_new_cases$DRate) - Weekly_new_cases$DRate) * 100

```
```{r}
ggplot(Weekly_new_cases) + geom_col(aes(x=Weekly,y=new_cases)) + 
  labs(title="Weekly new_cases",x="Date date", y="Weekly new_cases")

ggplot(Weekly_new_deaths) + geom_col(aes(x=Weekly,y=new_deaths)) + 
  labs(title="Weekly new_deaths",x="Date date", y="Weekly new_deaths") +
   ylim(0,1000)
``` 

#### Monthly new_cases and new_deaths

```{r}

Monthly_new_cases <- aggregate(new_cases~Monthly,US,FUN=sum)
Monthly_new_deaths <- aggregate(new_deaths~Monthly,US,FUN=sum)
Monthly_new_cases$DRate <- Monthly_new_deaths$new_deaths/Monthly_new_cases$new_cases
Monthly_new_cases$LivedSaved <- Monthly_new_cases$new_cases * (max(Monthly_new_cases$DRate) - Monthly_new_cases$DRate) * 100
```
```{r}
ggplot(Monthly_new_cases) + geom_col(aes(x=Monthly,y=new_cases)) +
  labs(title="Monthly new_cases")  +
  scale_y_continuous(labels=scales::comma)

ggplot(Monthly_new_deaths) + geom_col(aes(x=Monthly,y=new_deaths)) +
  labs(title="Monthly new_deaths")
```

#### New YORK COUNTY cASES AND DEATHS

```{r}
ggplot(CA) + geom_line(aes(x=date,y=new_cases)) +
  labs(title="New York City County Cases")
```

```{r}
ggplot(CA) + geom_line(aes(x=date,y=new_deaths)) +
  geom_smooth(aes(x=date,y=new_deaths),span=0.25) +
  labs(title="New York City County Deaths")
```
` 

### Total Cases and Deaths by County

```{r}
NYT <-NYT %>% group_by(county) %>% summarise(
      Cases=sum(new_cases),
      Deaths = sum(new_deaths))
```
